---
title: "Reproducible Research, Project 1"
author: "S. Agatoni"
date: "May, 2020"
output:
  github_document:
    keep_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='Figures/')
```


  
### 0. load required packages: 

```{r load}
if(!require("pacman")) install.packages(pacman); library(pacman)
pacman::p_load(ggplot2, dplyr, tidyr, chron, yaml)
```
 
  
  
### 1. Code for reading in the dataset and/or processing the data: 

```{r read}
data <- read.csv("activity.csv", header = TRUE)
data$date <- as.Date(data$date)
```

 
  
### 2. Histogram of the total number of steps taken each day: 

```{r hist}
daily_sum <- na.omit(data) %>% 
  group_by(date) %>% 
  summarise(total_steps = sum(steps))

ggplot(data = daily_sum) + 
  geom_histogram(aes(total_steps), fill = "cornflowerblue", color = "darkblue", bins = 5)
```



### 3. Mean and median number of steps taken each day: 

```{r mean}
avg <- as.integer(mean(daily_sum$total_steps))
median <- median(daily_sum$total_steps)

print(avg)
print(median)
```


### 4. Time series plot of the average number of steps taken (interval on x-axis): 

```{r time_series}
per_int <- na.omit(data) %>% 
  group_by(interval) %>% 
  summarise(average_steps = mean(steps))

ggplot(data = per_int, aes(x = interval, y = average_steps)) + 
  geom_line(color = "#FC4E07") 
```


  
### 5. The 5-minute interval that, on average, contains the maximum number of steps: 

```{r max_int}
max_int <- 
  per_int %>% 
  filter (average_steps == max(average_steps)) 

print(max_int$interval)
```


  
### 6.Replacing missing data (steps) with the average steps per interval: 

```{r missing}
total_NA <- sum(is.na(data))
print(total_NA)
```




```{r imputing NAs}

adj_data <- merge(data, per_int, by = "interval")

for (i in 1:nrow(adj_data)) {
  if (is.na(adj_data$steps[i])){
    adj_data$steps[i] <- adj_data$average_steps[i]
  }
}

```

 
 
```{r new mean}
adj_sum <- 
  adj_data %>% 
  group_by(date) %>% 
  summarise(total_steps = sum(steps))
  
new_avg <- as.integer(mean(adj_sum$total_steps))
new_median <- as.integer(median(adj_sum$total_steps))

print(new_avg)
print(new_median)
```


  
### 7. Histogram of the total number of steps taken each day after imputing missing values: 

```{r hist_2}
ggplot(data = adj_sum) + 
  geom_histogram(aes(total_steps), fill = "green2", color = "green4", bins = 5)
```

  
### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends: 

```{r grouping by is.weekend and interval}
adj_avg <- adj_data %>%
  mutate (weekend = is.weekend(date)) %>%
  group_by(weekend, interval) %>%
  summarise(avg_steps = mean(steps))
```
   
   
```{r panel_plot}
ggplot(data = adj_avg) + 
  geom_line(aes(x = interval, y = avg_steps)) + 
              facet_grid(rows = vars(weekend)) 
```
